---
title: "Visualise"
author: "Natasha Hardy & Ben Best"
date: "18/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About

Tutorial for data manipulation here: https://noaa-iea.github.io/r3-train/visualize.html

## 3.1 Read Data


Picking up with the table we downloaded last time (2.1.4 Read table read.csv()), let’s read the data directly from the URL and use readr’s read_csv():

```{r Read Data}

# libraries
library(here)
library(readr)
library(DT)


# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
# if ERDDAP server down (Error in download.file) with URL above, use this:
#    url_ac <- "https://raw.githubusercontent.com/noaa-iea/r3-train/master/data/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# download data
if (!file.exists(csv_ac))
  download.file(url_ac, csv_ac)

# read data
d_ac        <- read_csv(csv_ac, col_names = F, skip = 2)
names(d_ac) <- names(read_csv(csv_ac))

# show data
datatable(d_ac)

```

Note the use of functions in libraries here and readr that you may need to install from the Packages pane in RStudio.

There here::here() function starts the path based on looking for the *.Rproj file in the current working directory or higher level folder. In this case it should be the same folder as your current working directory so seems unnecessary, but it’s good practice for other situations in which you start running Rmarkdown files stored in subfolders (in which case the evaluating R Chunks assume the working directory of the .Rmd).

I prefer readr::read_csv() over read.csv() since columns of character type are not converted to type factor by default. It will also default to being read in as a tibble rather than just a data.frame.

## 3.2 Plot statically with ggplot2

## 3.2.1 Simple line plot + geom_line()

Let’s start with a simple line plot of total_fisheries_revenue_coastwide (y axis) over time (x axis) using the grammar of graphics principles by:

Feed the data as the first argument to ggplot().
Set up the aesthetics aes() as the second argument for specifying the dimensions of the plot (x and y).
Add (+) the geometry, or plot type.
From the Data Visualization with ggplot2 Cheatsheet (RStudio menu Help > Cheat Sheets), we have these aesthetics to plot based on the value being continuous


```{r Select Data}

library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)

```

```{r Simple line plot}

# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast

```


## 3.2.2 Trend line + geom_smooth()

Add a smooth layer based on a linear model (method = "lm").

```{r Add smooth layer}

p_coast + 
  geom_smooth(method = "lm")

```

Try changing the method argument by looking at the help documentation ?geom_smooth.


## 3.2.3 Distribution of values + geom_histogram()

What if you want to look at a distribution of the values? For instance, you might simulate future revenues by drawing from this distribution, in which case you would want to use geom_histogram().

```{r Data distribution}

d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram()

```


Try changing the binwidth parameter.


## 3.2.4 Series line plot aes(color = region)

Next, let’s also show the other regional values (CA, OR and WA; not coastwide) in the plot as a series with different colors. To do this, we’ll want to tidy the data into long format so we can have a column for total_fisheries_revenue and another region column to supply as the group and color aesthetics based on aesthetics we see are available for geom_line():

```{r Data manip fisheries revenue category}

library(stringr)
library(tidyr)

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot longer
  pivot_longer(-time) %>% 
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>% 
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn

```


## 3.2.5 Update labels + labs()

Next, let’s update the labels for the title, x and y axes, and the color legend:

```{r Edit labels}

p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn

```

